### 2008-09-15 Latex/H2Preview/H2PrevAbstract.tex
First submission

\label{sec:Abstract}
This paper is concerned with a broad class of $\htwo$ controller synthesis problems with preview. A generic preview design problem is 
embedded in a standard generalized regulator framework, and features both previewable and non-previewable disturbances. Preview regulation is acomplished by a two-degree-of-freedom output-feedback controller. The paper is concerned with a number of theoretical issues including the efficient solution of the standard $\htwo$ full-information Riccati equation and the efficient evaluation of the full-information preview gain matrices. The full-information problem is then extended to include the evaluation and efficient synthesis of the output-feedback controller.
 The synthesis of feedforward controllers with preview is analysed as a special case; this problem is of interest to designers who wish to introduce preview as a separate part of a system design. The way in which preview reduces the $\htwo$-norm of the closed-loop system is analyzed in detail.
Closed-loop norm reduction formulae provide a systematic way of establishing how much preview is required to solve a particular problem, and determine when extending the preview horizon will not produce worthwhile benefits. The paper concludes with a summary of the main features of preview control, as well as some controller design insights.