### 2008-09-15 Latex/H2Preview/H2PrevFIStandard.tex

%\label{subsec:StandardFI}
\subsection{Standard Theory}
\label{subsec:stdH2FI}
We begin with a brief summary of the discrete-time, linear time-invariant perfect information control problem \cite{LimebeerGreen}, which has plant description
$$P_{FI} \shorteq \arr{c}{\hspace{0.8cm}\stackrel{n}{\leftrightarrow}\hspace{0.5cm}\stackrel{l}{\leftrightarrow}\hspace{0.6cm}\stackrel{m}{\leftrightarrow} \\\arr{r}{\scriptstyle{n} \,\updownarrow\\\scriptstyle{p} \,\updownarrow\\\scriptstyle{n} \,\updownarrow\\\scriptstyle{l} \,\updownarrow} \ma{\arr{c|cc}{ A&B_1&B_2\\\hline C_1&D_{11}&D_{12}\\I&0&0\\0&I&0}}},$$and which satisfies the following standard assumptions:
\begin{itemize}
\item[(A1)] $(A,B_2)$ is stabilisable 
\item[(A2)] $D_{12}'D_{12}>0$
\item[(A3)] rank$\ma{A-e^{j\theta}I&B_2\\C_1& D_{12}}=n+m \quad \forall\, \theta \,\in (-\pi,\pi]$.
\end{itemize}

We would like to find the internally stabilising controller $K_{FI}$ which minimises $\nrm{F_l(P_{FI},K_{FI})}_2$. First define:
\aln{
\bar R&=D_{12}'D_{12}+B_2'XB_2   \\
  F_2&=-{\bar R}^{-1}(B_2'XA+D_{12}'C_1)  \label{eqn:F2Def}\\
A_c&=A+B_2F_2.
}

In \cite{ZDG} it is shown that if (A1)-(A3) are satisfied then there exists a solution $X$ to the Discrete Algebraic Riccati Equation (DARE):

\aln{X=A'XA-F_2'{\bar R}F_2+C_1'C_1 \label{eqn:X2d}}
such that:
\aln{
&X\geq 0 \label{eqn:Xpos}\\
&A_c \,\, \textrm{is asymptotically stable} \label{eqn:Xstab}
.}

A matrix $X$ satisfying (\ref{eqn:X2d}) and (\ref{eqn:Xstab}) is said to be \textit{stabilising}.

The internally stabilising, full information $\htwo$-optimal controller is then given by:
\begin{equation}
K_{FI}=\ma{F_2 &F_0} \label{eqn:FIcont}
\end{equation} 
with
\[
F_0=-{\bar R}^{-1}(B_2'XB_1+D_{12}'D_{11}).
\]
%
The resulting closed-loop norm is given by:
\mls{
\nrm{F_l(P_{FI},K_{FI})}_2^2=\\\tra{(D_{11}+D_{12}F_0)'(D_{11}+D_{12}F_0) +(B_1+B_2F_0)'X(B_1+B_2F_0) }.
}