### 2008-09-15 trunk/Latex/H2Preview/H2PrevConcs.tex

\label{sec:Conclusions}
Preview control has been studied for over three decades and many theoretical results can be found in the technical literature. Preview control has also been employed in a variety of contempory applications in mechanical engineering. These include active automotive suspension control \cite{Roh_1999_Stoc_Opt_Prev,Marzbanrad_2004_SuspPrev}, helicopter flight control \cite{Paulino_2006_PreviewRotorcraftAffine}, 
and driver steering control \cite{Cole_2006_PredictiveAndPreviewSteeringControl}. The purpose of the present work is to embed a class of preview control problems in a generalized regulator configuration \cite{LimebeerGreen,ZDG}. The open-loop system is controlled by a two-degree-of-freedom controller, which optimizes the system response to a combination of previewable and non-previewable exogenous inputs. Efficient algorithms have been presented for solving the two standard LQG Riccati equations, as well as computing the perfect information controller gains. The associated state estimator is low order with degree $n_g+n_r$. A detailed analysis of the achievable $\htwo$-norm cost reduction due to preview is also presented. The paper concludes with a summary of the main features of $\htwo$ preview control as well as some useful design insights.