### 2008-10-27 Latex/H2Preview/H2PrevAbstract.tex
resubmission 26/10/08

\label{sec:Abstract}
The purpose of this paper is to provide a set of synthesis and design tools for a wide class of $\htwo$ preview control systems. A generic preview design problem, which features both previewable and non-previewable disturbances,  is embedded in a standard generalized regulator framework. Preview regulation is accomplished by a two-degree-of-freedom output-feedback controller. A number of theoretical issues are studied, including the efficient solution of the standard $\htwo$ full-information Riccati equation and the efficient evaluation of the full-information preview gain matrices. The full-information problem is then extended to include the efficient implementation of the output-feedback controller. The synthesis of feedforward controllers with preview is analysed as a special case -- this problem is of interest to designers who wish to introduce preview as a separate part of a system design. The way in which preview reduces the $\htwo$-norm of the closed-loop system is analyzed in detail. Closed-loop norm reduction formulae provide a systematic way of establishing how much preview is required to solve a particular problem, and determine when extending the preview horizon will not produce worthwhile benefits. The paper concludes with a summary of the main features of preview control, as well as some controller design insights. New applications examples are introduced by reference.